---
title: "index"
author: "Jimena Velazquez de Leon"
date: "5/7/2021"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("network")
#install.packages("igraph")
library(readr)
library(dplyr)
library(tidyverse)
library(network)
library(igraph)
setwd("~/Desktop/BDS 516/git/jimevdl18")
```


Homework 10 

**Group Members:** Arvind Bala, Ben Brikman, Jimena Velazquez, Minglu Xu, Eva Zhang

# Question 1
We prepared the data for the network analysis by creating a separate edge and node list.

```{r, echo = F, include = F}
# Load data, create nodelist and edgelist

hw10_data <- read_csv("hw10_data.csv")
hw10_data = hw10_data[-1,]
hw10_data <- hw10_data[!is.na(hw10_data$Q2), ]

nodelist <- select(hw10_data,Q2,Q4) %>% 
  rename('ID' = Q2, 'Extraversion' = Q4) %>% 
  transform(ID = as.numeric(ID)) %>% 
  add_row(ID = 3, Extraversion = NA) %>% 
  add_row(ID = 4, Extraversion = NA) %>%
  add_row(ID = 8, Extraversion = NA) %>%
  add_row(ID = 9, Extraversion = NA) %>%
  add_row(ID = 10, Extraversion = NA) %>%
  add_row(ID = 11, Extraversion = NA) %>%
  add_row(ID = 12, Extraversion = NA) %>%
  add_row(ID = 13, Extraversion = NA) %>%
  add_row(ID = 14, Extraversion = NA) %>%
  add_row(ID = 16, Extraversion = NA) %>%
  add_row(ID = 19, Extraversion = NA) %>%
  add_row(ID = 21, Extraversion = NA) %>%
  add_row(ID = 22, Extraversion = NA) %>%
  add_row(ID = 25, Extraversion = NA) %>%
  add_row(ID = 26, Extraversion = NA) %>%
  add_row(ID = 27, Extraversion = NA) %>%
  add_row(ID = 29, Extraversion = NA) %>%
  add_row(ID = 37, Extraversion = NA) %>%
  add_row(ID = 39, Extraversion = NA) 

edgelist <- select(hw10_data,-Q4)
edgelist <- edgelist %>% 
  pivot_longer(!Q2, names_to = 'in-tie', values_to = 'tie-num', values_drop_na = TRUE) %>% 
  rename('out-tie' = Q2) %>% 
  separate('tie-num', into = c("1", "2","3","4","5"), sep = ",") %>% 
  pivot_longer('1':'5', names_to = 'number', values_to = 'tie-type', values_drop_na = TRUE) %>% 
  select('out-tie', 'in-tie', 'tie-type') 
edgelist$`in-tie` <- gsub("^.{0,3}", "", edgelist$`in-tie`)
```

# Question 2
*Calculate measures of centrality and similarity in networks depending on tie-type?
Which tie type generates the most similar network in terms of introversion? 
Which tie type is the most dissimilar?*

**Network Visualization by Tie Type**

We start our analysis by visualizing the overall network and the 5 networks
filtered by the 5 tie types. It's interesting to observe that
networks based on tie type 3, 4, 5 have relatively more isolated nodes. These ties 
describes friendships and relations of asking/being asked for help. Perhaps 
remote learning has impacted people's ability to make friends in the program.

```{r, echo = F}
detach(package:network)
newnetwork <- graph_from_data_frame(d = edgelist, vertices = nodelist, directed = TRUE)
# Plot overall network of all ties
plot(newnetwork, edge.arrow.size = 0.1, vertex.size = 3, main = "Overall Network")

# Plot overall network varying node size based on number of ties
#deg <- degree(newnetwork, mode="all")
#plot(newnetwork, vertex.size=deg/10, edge.arrow.size = 0.1, main = "Overall Network")

# filter by all 5 tie types
edgelist_1 <- edgelist %>% filter(`tie-type` == 1)
edgelist_2 <- edgelist %>% filter(`tie-type` == 2)
edgelist_3 <- edgelist %>% filter(`tie-type` == 3)
edgelist_4 <- edgelist %>% filter(`tie-type` == 4)
edgelist_5 <- edgelist %>% filter(`tie-type` == 5)

# Tie type 1 plot
newnetwork_1 <- graph_from_data_frame(d = edgelist_1, vertices = nodelist, directed = TRUE)
plot(newnetwork_1, edge.arrow.size = 0.1, vertex.size = 3, 
     main = "Tie Type 1 Network - Messaged Last Week")

# Tie type 2 plot
newnetwork_2 <- graph_from_data_frame(d = edgelist_2, vertices = nodelist, directed = TRUE)
plot(newnetwork_2, edge.arrow.size = 0.1, vertex.size = 3, 
     main = "Tie Type 2 Network - Met During MBDS")

# Tie type 3 plot
newnetwork_3 <- graph_from_data_frame(d = edgelist_3, vertices = nodelist, directed = TRUE)
plot(newnetwork_3, edge.arrow.size = 0.1, vertex.size = 3, 
     main = "Tie Type 3 Network - Is My Friend")

# Tie type 4 plot
newnetwork_4 <- graph_from_data_frame(d = edgelist_4, vertices = nodelist, directed = TRUE)
plot(newnetwork_4, edge.arrow.size = 0.1, vertex.size = 3, 
     main = "Tie Type 4 Network - I Ask for Help")

# Tie type 5 plot
newnetwork_5 <- graph_from_data_frame(d = edgelist_5, vertices = nodelist, directed = TRUE)
plot(newnetwork_5, edge.arrow.size = 0.1, vertex.size = 3, 
     main = "Tie Type 5 Network - Asks Me for Help")
```